---
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, overall-plots}
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(drlib)

theme_set(theme_minimal(20) +
            theme(legend.position = "bottom",
                  plot.title.position = "plot",
                  panel.grid.major.x = element_line(colour = "gray60"),
                  panel.grid.minor.x = element_blank(),
                  panel.grid.major.y = element_blank())
          )

files <- list.files(here("data"), full.names = TRUE)
d <- map_df(grep("A-F|B-F", files, value = TRUE), read_xlsx) %>% 
  clean_names() %>% 
  mutate(state = gsub(".+\\((.{2}).+", "\\1", where_tch))

regions <- read_xlsx(grep("DirectoryV3", files, value = TRUE), sheet = "Region")
  
regions <- regions %>% 
  mutate(state = gsub("\\('(.{2})'.+", 
                          "\\1", 
                          `State relation to region`)) %>% 
  select(state, region = Region)

d <- d %>% 
  left_join(regions) %>% 
  mutate(region = ifelse(where_tch == "International/non-US state or territory",
                         "International",
                         ifelse(!is.na(where_tch) & is.na(region),
                                "Territory",
                                region)))

d <- d %>% 
  mutate(yrs_teach = case_when(yrs_teach == "43529" ~ "3-5",
                               yrs_teach == "43626" ~ "6-10",
                               yrs_teach == "43784" ~ "11-15",
                               TRUE ~ yrs_teach))
grade_selection <- d %>% 
  select(rspdnt_id, pre_k, gk, matches("g\\d?\\d")) %>% 
  gather(var, grade, - rspdnt_id) %>% 
  drop_na(grade) 

d <- semi_join(d, 
          filter(grade_selection, grade == "K"))
```

# Background and System Use/Experience

## How did you learn about ReadWorks?
```{r, fig.width = 9}
learn_methods <- d %>% 
  select(rspdnt_id, starts_with("learn")) %>% 
  gather(option, response, -rspdnt_id) %>% 
  drop_na(response) %>% 
  count(response) %>% 
  mutate(response = reorder(response, n))

ggplot(learn_methods, aes(response, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## How often do you visit/use ReadWorks
```{r, fig.width = 10}
d %>% 
  select(rspdnt_id, visit_rw, use_rw) %>% 
  gather(what, rate, -rspdnt_id) %>% 
  count(rate, what) %>%
  mutate(what = ifelse(what == "use_rw", "Use", "Visit"),
         rate = reorder(rate, n)) %>% 
ggplot(aes(rate, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  facet_wrap(~what) +
  labs(y = "Number of Responses",
       x = "")
```

## Would you recommend ReadWorks? {.tabset .tabset-fade}

### Overall Scores

```{r}
ggplot(d, aes(rate1_10)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

### Why did you choose the score you did?

```{r by_score}
by_score <- d %>% 
  select(rspdnt_id, rate1_10, matches("^r\\d")) %>% 
  gather(var, response, matches("^r\\d")) %>% 
  drop_na(response) %>% 
  count(rate1_10, response) %>% 
  nest(-rate1_10) 

by_score <- by_score %>% 
  mutate(data = map(data, ~mutate(.x ,response = reorder(response, n))),
         plot = map2(data, rate1_10, ~
           ggplot(.x, aes(response, n)) +
           geom_col(fill = "#F4E8C1", alpha = 0.9) +
           coord_flip() +
           ylim(0, max(by_score$data[[nrow(by_score)]]$n)) +
           labs(y = "Number of Responses",
                x = "",
                title = paste("Why did you choose a score of", .y)) +
          theme_minimal(base_size = 10)
                      )
         ) 

walk(rev(by_score$plot), print)
```

## {-}

## How much of your ELA block is used for literacy?

```{r}
d %>% 
  drop_na(amt_ela_lit_ctr_percent) %>% 
  mutate(ela_percent = factor(amt_ela_lit_ctr_percent,
                              levels = c("None", 
                                         "About 10%",
                                         "About 25%",
                                         "About 50%",
                                         "About 75%",
                                         "Greater than 75%"))) %>% 
ggplot(aes(ela_percent)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")

```

## How do you use ReadWorks with students in different configurations?
```{r, fig.height = 25, fig.width = 10}
new_labs <- d %>% 
  select(rspdnt_id, amt_ela_lit_ctr_percent, matches("^ar\\d")) %>% 
  gather(var, response, starts_with("ar")) %>% 
  drop_na(response, amt_ela_lit_ctr_percent) %>% 
  count(var, response) %>%
  mutate(new_lab = c(
    "I do not use Readworks in stations",
    "Independent work with follow-up",
    "Independent work with teacher follow-up",
    "Small groups with follow-up",
    "Small groups with teacher follow-up",
    "Work with individual students",
    "Small groups"
  )) %>% 
  select(-n)



d %>% 
  select(rspdnt_id, amt_ela_lit_ctr_percent, matches("^ar\\d")) %>% 
  gather(var, response, starts_with("ar")) %>% 
  drop_na(response, amt_ela_lit_ctr_percent) %>% 
  count(response, amt_ela_lit_ctr_percent) %>% 
  left_join(new_labs) %>% 
ggplot(aes(new_lab, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  facet_wrap(~amt_ela_lit_ctr_percent, ncol = 1) +
    labs(y = "Number of Responses",
         x = "",
         title = "Plots separated by proportion of time spent in centers")
```

## What are the main ways you use ReadWorks?
```{r, fig.width = 12}
d %>% 
  select(matches("device\\d")) %>% 
  gather(device, response) %>% 
  drop_na(response) %>% 
  count(response) %>% 
  mutate(response = c(
           "I do not use ReadWorks texts",
           "Aligning with other curricula",
           "Aligning with ELA curricula",
           "Aligning with science/social studies curricula",
           "Matching to my students' reading levels",
           "Supporting an activity"
         ),
         response = reorder(response, n)) %>% 
ggplot(aes(response, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "",
       "I mostly think about...")
```


## Do you use ReadWorks vocabulary supports and/or activities? {.tabset .tabset-fade}

### Overall

```{r}
d %>% 
  select(rw_voc_use_yn) %>% 
  drop_na(rw_voc_use_yn) %>% 
ggplot(aes(rw_voc_use_yn)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

### Why?

```{r fig.width = 15, fig.height = 10}
d %>% 
  select(starts_with("voc_r")) %>% 
  gather(var, response) %>% 
  drop_na(response)  %>%
  count(response) %>% 
  mutate(response = reorder(response, n)) %>% 
ggplot(aes(response, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "",
       title = "Why do you use ReadWorks vocabulary supports?")
```

### Why not?
```{r fig.width = 15, fig.height = 10}
d %>% 
  select(starts_with("voc_n")) %>% 
  gather(var, response) %>% 
  drop_na(response)  %>%
  count(response) %>% 
  mutate(response = reorder(response, n)) %>% 
  ggplot(aes(response, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "",
       title = "Why do you NOT use ReadWorks vocabulary supports?")
```

## {-}

## Does ReadWorks improve reading skills?
```{r, fig.height = 25}
new_labs <- tibble(
  var = c("bck_know", "build_voc", "chlg_txt", "excite", "imp_comp", 
          "rec_t_struc"),
  lab = c("Build background knowledge?", "Build vocabulary?", 
          "Read more challenging text?", "Build reading excitement?",
          "Improve comprehension?", "Recognize text structure?")
)

d %>%
  select(excite, imp_comp, bck_know, build_voc, chlg_txt, rec_t_struc) %>% 
  gather(var, val) %>% 
  drop_na(val) %>% 
  count(var, val) %>% 
  left_join(new_labs) %>% 
  mutate(val = factor(val, 
                      levels = c("I don't know", "None at all", "A little",
                                 "A moderate amount", "A lot", "A great deal")
                      )) %>% 
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  facet_wrap(~lab, ncol = 1) +
  labs(y = "Number of Responses",
       x = "")
```


## Do you have regular access to a projector smartboard?

```{r}
d %>%
  drop_na(access_sb) %>%
ggplot(aes(access_sb)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```


## How many devices are in the classroom
```{r, fig.height = 12, fig.width = 15}
d %>%
  drop_na(num_devices) %>%
  mutate(num_devices = factor(
    num_devices,
    levels = c("None",
               "1 for me to use for instruction and none for students to use",
               "1 for every 10 students or more",
               "1 for every 5-9 students",
               "1 for every 2-4 students",
               "1 for every student")
      )
    ) %>%
ggplot(aes(num_devices)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "") +
  theme(axis.text.x = element_text(size = 25),
        axis.text.y = element_text(size = 25),
        axis.title.x = element_text(size = 35))
```

## What devices do you use?
```{r, fig.width = 10}
d %>%
  select(dev_access, tech_use1, tablet1, smart_ph1) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(var, val) %>%
  mutate(val = reorder(val, n)) %>%
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## Which devices do students use?
```{r, fig.width = 12}
d %>%
  select(which_dev, tech_use2, tablet2, smartph2) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(var, val) %>%
  mutate(val = reorder(val, n)) %>%
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## Which devices do students use outside of school?
```{r, fig.width = 10}
d %>%
  select(starts_with("device_ac")) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(var, val) %>%
  mutate(val = ifelse(val == "I don’t know", "I don't know", val),
         val = factor(
           val,
           levels = c(
             "My students do not access edtech resources outside of school",
             "I don't know",
             "Smartphones",
             "Tablets",
             "Desktop/laptop computers (including Chromebooks)"
             )
           )
         ) %>%
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## How easy is it to grade written assignments?
```{r fig.width = 10}
d %>%
  drop_na(ease_grade_res) %>%
  mutate(ease_grade_res =
           factor(ease_grade_res,
                  levels = c("I have never graded student written responses within the ReadWorks digital platform",
                             "Very easy", "Easy", "Hard", "Very hard"),
                  labels = c("No experience", "Very easy", "Easy", "Hard", "Very hard"))) %>%
  ggplot(aes(ease_grade_res)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## How easy is it to find materials
```{r fig.width = 10}
d %>%
  drop_na(find_mat_rw) %>%
  mutate(find_mat_rw =
           factor(find_mat_rw,
                  levels = c("I have never searched the ReadWorks digital platform for materials or supports for individual students",
                             "Very easy", "Easy", "Hard", "Very hard"),
                  labels = c("No experience",
                             "Very easy", "Easy", "Hard", "Very hard"))) %>%
ggplot(aes(find_mat_rw)) +
  geom_bar(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What troubles do you have?
```{r, fig.height = 10}
d %>%
  select(starts_with("why")) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(var, val) %>%
  mutate(val = c("None", "Finding time to learn system",
                 "Difficult to understand","Takes too much time",
                 "Don't have devices", "Old devices",
                 "No access outside of school", "Time on devices",
                 "Prefer paper"),
         val = reorder(val, n)) %>%
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What new features would you like to see?
```{r, fig.width = 15, fig.height = 25}
d %>%
  select(starts_with("feat")) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(var, val) %>%
  mutate(val = reorder(val, n)) %>%
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  scale_y_continuous(expand =  c(0, 0)) +
  labs(y = "Number of Responses",
       x = "") +
  theme(axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20),
        axis.title.x = element_text(size = 30))
```

## Variables within instructional applications using ReadWorks

```{r, fig.height = 25}
percent_var_labs <- data.frame(
  var = c("bac_know_percent", "comp_strat_percent", "indep_read_percent",
          "morph_wf_percent", "read_aloud_percent", "test_take_percent",
          "voc_defs_percent"),
  lab = c("Background Knowledge", "Comprehension Strategies",
          "Independent Reading", "Morpheme", "Read aloud", "Test Taking",
          "Vocabulary")
)

percents_count <- d %>%
  select(bac_know_percent, comp_strat_percent, voc_defs_percent,
         morph_wf_percent, test_take_percent, indep_read_percent,
         read_aloud_percent) %>%
  gather(var, percent) %>%
  drop_na(percent) %>%
  left_join(percent_var_labs) %>%
  count(lab, percent)

percents_count$percent <- factor(
  percents_count$percent,
  levels = unique(percents_count$percent)[c(6:7, 1:4, 9, 5, 8)],
  labels = unique(percents_count$percent)[c(6:7, 1:4, 4, 8, 8)]
)

ggplot(percents_count, aes(percent, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  facet_wrap(~lab, ncol = 1) +
  labs(y = "Number of Responses",
       x = "")
```

## Trust-related variables
```{r fig.height = 15, fig.width = 8}
d %>%
  select(effect_tchr_yn, trust_rwlm, qual_rwlm, rw_qt) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(var, val) %>%
  mutate(val = ifelse(nchar(val) > 30, "Never used", val),
         val = ifelse(val == "Higher quality", "Higher Quality", ifelse(
                      val == "Lesser quality", "Lesser Quality", val)),
         var = case_when(var == "effect_tchr_yn" ~
                           "Increase teaching effectiveness?",
                         var == "qual_rwlm" ~
                           "Higher quality than other materials?",
                         var == "rw_qt" ~
                           "Do you trust ReadWorks for guidance/training?",
                         TRUE ~
                           "Do you trust ReadWorks more/less than other materials?"
                         ) ) %>%
ggplot(aes(val, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  coord_flip() +
  facet_wrap(~var, scales = "free_y", ncol = 1) +
  labs(y = "Number of Responses",
       x = "")
```

# Demographics
## What's your role?
```{r fig.height = 8}
d %>%
  count(role) %>%
  drop_na(role) %>%
  mutate(role = reorder(role, n)) %>%
  ggplot(aes(role, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What's your school program?
```{r}
d %>%
  count(school_prog) %>%
  drop_na(school_prog) %>%
  mutate(school_prog = reorder(school_prog, n)) %>%
ggplot(aes(school_prog, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## How many years have you taught?
```{r}
d %>%
  count(yrs_teach) %>%
  drop_na(yrs_teach) %>%
  mutate(yrs_teach = reorder(yrs_teach, n)) %>%
ggplot(aes(yrs_teach, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## Do you teach in a Title 1 school?
```{r}
d %>%
  count(tch_title) %>%
  drop_na(tch_title) %>%
  mutate(tch_title = reorder(tch_title, n)) %>%
ggplot(aes(tch_title, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What grades do you teach?
```{r, fig.height = 8}
grades_taught <- d %>%
  select(pre_k, gk, matches("^g\\d?\\d$"), post_sec, adults, other_grade) %>%
  gather(grade, selection) %>%
  drop_na(selection) %>%
  count(selection) %>%
  mutate(selection = factor(selection,
                            levels = rev(c("Pre-school/PK", "K", 1:12,
                                       "Post-secondary", "Adults", "Other"))))

ggplot(grades_taught, aes(selection, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What content area do you teach?

```{r fig.height = 7.75}
content <- d %>%
  select(ela = subjects, math, science, soc_stud, world_lang, art_mus, library,
         tech_cs, other_content) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(val) %>%
  mutate(content = reorder(val, n))

ggplot(content, aes(content, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What population do you currently teach?
```{r fig.height = 8, fig.width = 10}
stu_pops <- d %>%
  select(gened = pops_teach, sped, ell, dyslex, trauma, low_inc) %>%
  gather(var, val) %>%
  drop_na(val) %>%
  count(val) %>%
  mutate(pop = reorder(val, n))

ggplot(stu_pops, aes(pop, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## What non-English language do your students speak
````{r fig.height = 9, fig.width = 9}
languages <- d %>%
  select(lang_non_eng, arabic, chinese, french, german, hatian, korean,
         russian, spanish, tagalog, vietnamese, other_lang) %>%
  gather(var, val) %>%
  mutate(val = ifelse(val == "I don’t know", "I don't know", val)) %>%
  drop_na(val) %>%
  count(val) %>%
  mutate(language = reorder(val, n))

ggplot(languages, aes(language, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```

## Where do you teach
```{r fig.height = 20, fig.width = 9}
d %>%
  count(where_tch) %>%
  drop_na(where_tch) %>%
  mutate(where_tch = gsub("(.+) \\(.+", "\\1", where_tch),
         where_tch = reorder(where_tch, n)) %>%
ggplot(aes(where_tch, n)) +
  geom_col(fill = "#F4E8C1", alpha = 0.9) +
  scale_y_continuous(expand =  c(0, 0)) +
  coord_flip() +
  labs(y = "Number of Responses",
       x = "")
```
